Case Study (McKenna)

### 2020-11-20 Cases/mrp/Case Study (McKenna) .Rmd
complete updates

In this case study there are five classification variables: age, income, ethnicity, gender, and state. Each variable is transformed into categorical variables with age having seven categories, income having three, ethnicity having three, gender having two, and state have 50 making the total number of cells 6,300 (7 x 3 x 3 x 2 x 50 = 6300).

@@

```{r, eval = FALSE}
options(mc.cores = parallel::detectCores()) # Set Stan to use all available cores.
rstan_options(auto_write = TRUE)            # Don't recompile Stan code that hasn't changed.

# Data generating function for reference. 
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) 
  # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) 
#Adjust the N to be the number observed in each category in each group
  }
# Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)
  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }
  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }
#male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]
  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))
  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r}
# Importing the data, will see it exported in environment 
sample <- read_csv(here::here("Data", "sample.csv"))
poststrat <- read_csv(here::here( "Data", "poststrat.csv"))
true_popn <- read_csv(here::here( "Data", "true_popn.csv"))

# Add groups to data. 
poststrat_group <- poststrat %>% 
  mutate(group = row_number())

sample_group <- left_join(sample,
                          poststrat_group,
                          by = c(
                            "male" = "male", 
                            "age" = "age", 
                            "eth" = "eth", 
                            "income" = "income",
                            "state" = "state") )
```

@@

```{r}
predictors <- sample_group %>% 
  select(male, age, eth, income, state)

data_list <- list(
  D = 5,                                              # Number of variables. 
  N = 1200,                                           # Number of observations.
  L = 6300,                                           # Number of groups. 
  y = sample_group$service_failure,                   # Outcome variables. 
  ll = sample_group$group,                            # Group assignment. 
  x = predictors                                      # Matrix of predictors. 
)
```

>>>>

In this case study there are three classification variables: age, ethnicity, and gender. Each variable is transformed into categorical variables with age having seven categories, ethnicity having three, and gender having two making the total number of cells 42 (7 x 3 x 2 = 42).

@@

```{r}
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7) # male or not, eth, age
  # poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) 
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)))) 
  # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age")
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
            count <- count + 1
            # Fill them in so we know what category we are referring to
            poststrat[count, 1:3] <- c(i1-1, i2, i3)
          }
        }
      }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.2, 0.3, 0.5)
  p_age <- c(0.1, 0.1, 0.2, 0.2, 0.1, 0.1, 0.2)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
                              p_age[poststrat[j,3]]) 
    #Adjust the N to be the number observed in each category in each group
  }
  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)
  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }
  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:3], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
    )
  }
  #male or not, eth, age
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  sample <- data.frame(service_failure = y,
                       male, age, eth,
                       id = 1:length(people))
  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
mrp_sim <- simulate_mrp_data(n=1200)
sample <- mrp_sim[["sample"]]
poststrat <- mrp_sim[["poststrat"]]
true_popn <- mrp_sim[["true_popn"]]
```

```{r}
# Add groups to data. 
poststrat_group <- poststrat %>% 
  mutate(group = row_number())

sample_group <- left_join(sample,
                          poststrat_group,
                          by = c(
                            "male" = "male", 
                            "age" = "age", 
                            "eth" = "eth") )
```

@@

```{r, eval= FALSE}
predictors <- sample_group %>% 
  select(male, age, eth)

data_list <- list(
  D = 3,                                              # Number of variables. 
  N = 1200,                                           # Number of observations.
  L = 42,                                             # Number of groups. 
  y = sample_group$service_failure,                   # Outcome variables. 
  ll = sample_group$group,                            # Group assignment. 
  x = predictors                                      # Matrix of predictors. 
)
```
